---
title: "Microbiome Timeseries"
author: "Tess Attenborough"
date: "30 April 2019"
output: html_document

# Load libraries
```{r}
library(data.table)
library(zoo)
library(plyr)
library(scales)
library(ggplot2)
library(ggthemes)
library(dplyr)
library(plotly)
library(pracma)
library(Hmisc)
library(rpart)
library(survminer)
library(knitr)
library(grid)
library(stringr)
library(ggrepel)
library(ggjoy)
library(RColorBrewer)
library(stringdist)
library(ape)
library(reshape2)
library(gplots)
library(ineq)
library(vegan)
library(Rtsne)
library(praise)
```

Next, a multiplot function to produce ggplot grid plots.
```{r}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

# Load data into format suitable for analysis

```{r}
#set the working directory
wd <- "/Users/teresaattenborough/CVGSTUDY/CVGR"
wd.output <- "/Users/teresaattenborough/CVGSTUDY/CVGR/TimeSeries"
setwd(wd)

CVG017Freqs <- read.csv(file = "CVG017.csv", stringsAsFactors = F) #import otu table

CVG017Freqs <- CVG017Freqs[!(CVG017Freqs$Frequency==0),]


#import a dataframe suitable (should contain these things)
#dat <- data.frame(OTus, Days_pst, freq,  stringsAsFactors = F)
#CD4
#ID, month, chain, CDR3, freq
Days_post <- CVG017Freqs$Days_post_transplant #days post transplant
OTUs<- CVG017Freqs$OTU #OTU
Count <- CVG017Freqs$Frequency #Freqs
ID <- "CVG017"
Genus <- CVG017Freqs$Genus



dat <- data.frame(ID, OTUs, Count, Days_post, Genus, stringsAsFactors = F)

```


# % shared clonotypes time
```{r}
#  y | x1 | x2 | shared |
# ---|----|----|--------|
#  1 |  1 |  2 |   50   |
#  1 |  2 |  3 |   50   |
#  1 |  1 |  3 |   25   |

fillseq <- seq(1, length(CVG017dfm$OTU)) #make list of numbers corresponding with number of unique clonotypes

write.csv(CVG017dfm, file = "CVG017dfm.csv")

#here i remove from the dataframe any OTUs that were either classed as contmainants or were not able to be classified
CVG017dfm <- subset(CVG017dfm, OTU != "OTU_180" & OTU != "OTU_268" & OTU != "OTU_344" & OTU != "OTU_354" & OTU != "OTU_360" & OTU != "OTU_372" & OTU != "OTU_117" & OTU != "OTU_137" & OTU != "OTU_30" & OTU != "OTU_19" & OTU != "OTU_129" & OTU != "OTU_156" & OTU != "OTU_188" & OTU != "OTU_260" & OTU != "OTU_76" & OTU != "OTU_241" & OTU != "OTU_242" & OTU != "OTU_63" & OTU != "OTU_106" & OTU != "OTU_135" & OTU != "OTU_303" & OTU != "OTU_222" & OTU != "OTU_56" & OTU != "OTU_120" & OTU != "OTU_133" & OTU != "OTU_200" & OTU != "OTU_146" & OTU != "OTU_177" & OTU != "OTU_507" & OTU != "OTU_247" & OTU != "OTU_249")

fillseq <- seq(1, length(CVG017dfm$OTU)) #make list of numbers corresponding with number of unique clonotypes

CVG017dfm <- CVG017dfm[order(CVG017dfm$Genus),]

CVG017dfm$fillseq <- fillseq #add numbers from fillseq to data frame with unique clonotypes



#time point 1
ID <- "CVG017"
OTUs<- CVG017dfm$OTU
Count <- CVG017dfm$CVG14V54122
Days_post <- "-6"
Genus <- CVG017dfm$Genus
fillseq <- CVG017dfm$fillseq
temp1 <- data.frame(ID, OTUs, Count, Days_post, Genus, fillseq, stringsAsFactors = F) #t0 first time point
temp1 <- temp1[!(temp1$Count==0),]

#time point 2
ID <- "CVG017"
OTUs<- CVG017dfm$OTU
Count <- CVG017dfm$CVG14V54156
Days_post <- "1"
Genus <- CVG017dfm$Genus
fillseq <- CVG017dfm$fillseq
temp2 <- data.frame(ID, OTUs, Count, Days_post, Genus, fillseq, stringsAsFactors = F)
temp2 <- temp2[!(temp2$Count==0),] #t2 the second time point   

#time point 3
ID <- "CVG017"
OTUs<- CVG017dfm$OTU
Count <- CVG017dfm$CVG14V54194
Days_post <- "35"
Genus <- CVG017dfm$Genus
fillseq <- CVG017dfm$fillseq
temp3 <- data.frame(ID, OTUs, Count, Days_post, Genus, fillseq, stringsAsFactors = F)
temp3 <- temp3[!(temp3$Count==0),] #t3 the third time point   

#time point 4
ID <- "CVG017"
OTUs<- CVG017dfm$OTU
Count <- CVG017dfm$CVG14V54331
Days_post <- "147"
Genus <- CVG017dfm$Genus
fillseq <- CVG017dfm$fillseq
temp4 <- data.frame(ID, OTUs, Count, Days_post, Genus, fillseq, stringsAsFactors = F)
temp4 <- temp4[!(temp4$Count==0),]   
 #t4 the fourth time point  

#time point 5
ID <- "CVG017"
OTUs<- CVG017dfm$OTU
Count <- CVG017dfm$CVG14V54407
Days_post <- "180"
Genus <- CVG017dfm$Genus
fillseq <- CVG017dfm$fillseq
temp5 <- data.frame(ID, OTUs, Count, Days_post, Genus, fillseq, stringsAsFactors = F)
temp5 <- temp5[!(temp5$Count==0),]   
 #t5 the fifth time point 


        ### Shared clonotype plot ###
        # creates a vector identifying shared sequences
temp1$shared <- NA #creates column
temp2$shared <- NA
temp3$shared <- NA #creates column
temp4$shared <- NA
temp5$shared <- NA


temp1$shared[temp1$OTUs %in% temp2$OTUs | temp1$OTUs %in% temp3$OTUs | temp1$OTUs %in% temp4$OTUs | temp1$OTUs %in% temp5$OTUs] <- T # marks clonotypes that are in temp1 and also in temp2 OR temp 3 OR temp4 as TRUE
temp1$shared[!temp1$OTUs %in% temp2$OTUs & !temp1$OTUs %in% temp3$OTUs & !temp1$OTUs %in% temp4$OTUs & !temp1$OTUs %in% temp5$OTUs] <- F # marks clonotypes that are in temp1 and NOT in temp2 OR temp3 OR temp4 as FALSE

temp2$shared[temp2$OTUs %in% temp1$OTUs | temp2$OTUs %in% temp3$OTUs | temp2$OTUs %in% temp4$OTUs | temp2$OTUs %in% temp5$OTUs] <- T  # marks clonotypes that are in temp2 and temp1 as TRUE
temp2$shared[!temp2$OTUs %in% temp1$OTUs & !temp2$OTUs %in% temp3$OTUs & !temp2$OTUs %in% temp4$OTUs & !temp2$OTUs %in% temp5$OTUs] <- F#  marks clonotypes that are in temp2 and NOT in temp1 as FALSE

temp3$shared[temp3$OTUs %in% temp1$OTUs | temp3$OTUs %in% temp2$OTUs | temp3$OTUs %in% temp4$OTUs | temp3$OTUs %in% temp5$OTUs] <- T  # marks clonotypes that are in temp2 and temp1 as TRUE
temp3$shared[!temp3$OTUs %in% temp1$OTUs & !temp3$OTUs %in% temp2$OTUs & !temp3$OTUs %in% temp4$OTUs & !temp3$OTUs %in% temp5$OTUs] <- F#  marks clonotypes that are in temp2 and NOT in temp1 as FALSE

temp4$shared[temp4$OTUs %in% temp1$OTUs | temp4$OTUs %in% temp2$OTUs | temp4$OTUs %in% temp3$OTUs | temp4$OTUs %in% temp5$OTUs] <- T  # marks clonotypes that are in temp2 and temp1 as TRUE
temp4$shared[!temp4$OTUs %in% temp1$OTUs & !temp4$OTUs %in% temp2$OTUs & !temp4$OTUs %in% temp3$OTUs & !temp4$OTUs %in% temp5$OTUs] <- F#  marks clonotypes that are in temp2 and NOT in temp1 as FALSE
   
temp5$shared[temp5$OTUs %in% temp1$OTUs | temp5$OTUs %in% temp2$OTUs | temp5$OTUs %in% temp3$OTUs | temp5$OTUs %in% temp4$OTUs] <- T  # marks clonotypes that are in temp2 and temp1 as TRUE
temp5$shared[!temp5$OTUs %in% temp1$OTUs & !temp5$OTUs %in% temp2$OTUs & !temp5$OTUs %in% temp3$OTUs & !temp5$OTUs %in% temp4$OTUs] <- F#  marks clonotypes that are in temp2 and NOT in temp1 as FALSE

#testing to see if this the way
#temp1$sharedintersect <- NA #creates column
#temp2$sharedintersect <- NA
#temp3$sharedintersect <- NA #creates column
#temp4$sharedintersect <- NA

#temp1$sharedintersect[temp1$V1 %in% temp2$V1] <- T
#temp1$sharedintersect[!temp1$V1 %in% temp2$V1] <- F #same for false
#temp2$sharedintersect[temp2$V1 %in% temp1$V1] <- T
#temp2$sharedintersect[!temp2$V1 %in% temp1$V1] <- F


temp1 <- arrange(temp1,desc(temp1$shared),temp1$fillseq) #orders the file first by seqs that are shared or not, then sorts by fillseq
temp2 <- arrange(temp2,desc(temp2$shared),temp2$fillseq)
temp3 <- arrange(temp3,desc(temp3$shared),temp3$fillseq)
temp4 <- arrange(temp4,desc(temp4$shared),temp4$fillseq)
temp5 <- arrange(temp5,desc(temp5$shared),temp5$fillseq)


  # vector to order and colour sequence
        #temp1$fillseq[temp1$shared == T] <- fillseq
        #temp1$fillseq[temp1$shared == F] <- length(temp1$fillseq[temp1$shared == T])+1
       # temp2$fillseq <- temp1$fillseq[match(temp2$V1, temp1$V1)]
        #temp2$fillseq[temp2$shared == F] <- length(temp2$fillseq[temp2$shared == T])+1
               

        # create 1 entry for all non-shared sequences
temp1.mod <- temp1 #duplicates temp1 and saves it as temp1.mod
temp2.mod <- temp2
temp3.mod <- temp3 #duplicates temp3 and saves it as temp3.mod
temp4.mod <- temp4
temp5.mod <- temp5



#temp3.mod$Count[length(temp3.mod$ID)] <- sum(temp3.mod$Count[temp3.mod$shared == F]) # the last entry is given the frequency of all the non-shared clonotypes
temp5.mod$Count[length(temp5.mod$ID)] <- sum(temp5.mod$Count[temp5.mod$shared == F])


#temp1.mod$V2[length(temp1.mod$ID)] <- sum(temp1.mod$V2[temp1.mod$sharedintersect == F]) # the last entry is given the frequency of all the non-shared clonotypes
#temp2.mod$V2[length(temp2.mod$ID)] <- sum(temp2.mod$V2[temp2.mod$sharedintersect == F])


#temp3.mod$shared[length(temp3.mod$ID)] <- T #makes the last entry true so it is kept
temp5.mod$shared[length(temp5.mod$ID)] <- T


#temp1.mod$sharedintersect[length(temp1.mod$ID)] <- T #makes the last entry true so it is kept
#temp2.mod$sharedintersect[length(temp2.mod$ID)] <- T


temp3.mod <- temp3.mod[!temp3.mod$shared == F,] #gets rid of all non-shared clonotypes
temp5.mod <- temp5.mod[!temp5.mod$shared == F,]

#temp1.mod <- temp1.mod[!temp1.mod$sharedintersect == F,] #gets rid of all non-shared clonotypes
#temp2.mod <- temp2.mod[!temp2.mod$sharedintersect == F,]


temp5.mod$shared[length(temp5.mod$ID)] <- F



#temp3.mod$fillseq[length(temp3.mod$ID)] <- 900000 #puts the last entry back last in fillseq order
temp5.mod$fillseq[length(temp5.mod$ID)] <- 900000

        
        # normalise sequence frequency
temp1.mod$Count <- (100/sum(temp1.mod$Count))*temp1.mod$Count #turns counts into percentages
temp2.mod$Count <- (100/sum(temp2.mod$Count))*temp2.mod$Count
temp3.mod$Count <- (100/sum(temp3.mod$Count))*temp3.mod$Count #turns counts into percentages
temp4.mod$Count <- (100/sum(temp4.mod$Count))*temp4.mod$Count
temp5.mod$Count <- (100/sum(temp5.mod$Count))*temp5.mod$Count

        
        #temp1.mod.expanded <- temp1.mod[rep(row.names(temp1.mod), temp1.mod$V2),]
        #temp2.mod.expanded <- temp2.mod[rep(row.names(temp2.mod), temp2.mod$V2),]
        
maxPcntShared <- max(c(100 - sum(temp1.mod$Count[temp1.mod$shared==T]), 100 - sum(temp2.mod$Count[temp2.mod$shared==T]), 100 - sum(temp3.mod$Count[temp3.mod$shared==T]), 100 - sum(temp4.mod$Count[temp4.mod$shared==T]), 100 - sum(temp5.mod$Count[temp5.mod$shared==T]))) #calculates highest % of uniques

minPcntShared <- min(c(100 - sum(temp1.mod$Count[temp1.mod$shared==T]), 100 - sum(temp2.mod$Count[temp2.mod$shared==T]), 100 - sum(temp3.mod$Count[temp3.mod$shared==T]), 100 - sum(temp4.mod$Count[temp4.mod$shared==T]), 100 - sum(temp5.mod$Count[temp5.mod$shared==T]))) #calculates lowest % of uniques (gives % of reads that come from clonotypes unique to the sample)
        
        # create x coordinates for line plotting
#temp1.mod$x <- 1.45 #new column all equalling 1.45
#temp2.mod$x <- 2.55 #new column all equalling 2.55
#temp3.mod$x <- 3.65 #new column all equalling 3.65
#temp4.mod$x <- 4.75 #new column all equalling 4.75

        # create x coordinates for line plotting
temp1.mod$x <- 5.85 #new column all equalling 1.45
temp2.mod$x <- 4.75 #new column all equalling 2.55
temp3.mod$x <- 3.65 #new column all equalling 3.65
temp4.mod$x <- 2.55 #new column all equalling 4.75
temp5.mod$x <- 1.45 #new column all equalling 4.75

        
        # create the y point for the plot
temp1.mod$y <- 100 - cumsum(temp1.mod$Count) + temp1.mod$Count #calculates y coordinate for each clonotype
temp2.mod$y <- 100 - cumsum(temp2.mod$Count) + temp2.mod$Count
temp3.mod$y <- 100 - cumsum(temp3.mod$Count) + temp3.mod$Count #calculates y coordinate for each clonotype
temp4.mod$y <- 100 - cumsum(temp4.mod$Count) + temp4.mod$Count
temp5.mod$y <- 100 - cumsum(temp5.mod$Count) + temp5.mod$Count

    
#collect relevant data
intersectA1 <- temp1.mod[temp1.mod$OTUs %in% temp2.mod$OTUs,] # overlap overlap( finds all the seqs in temp1 that also appear in temp 2)
intersectA2 <- temp2.mod[temp2.mod$OTUs %in% temp1.mod$OTUs,] # overlap overlap( finds all the seqs in temp1 that also appear in temp 2)
intersectB1 <- temp2.mod[temp2.mod$OTUs %in% temp3.mod$OTUs,] # overlap overlap( finds all the seqs in temp1 that also appear in temp 2)
intersectB2 <- temp3.mod[temp3.mod$OTUs %in% temp2.mod$OTUs,] # overlap overlap( finds all the seqs in temp1 that also appear in temp 2)
intersectC1 <- temp3.mod[temp3.mod$OTUs %in% temp4.mod$OTUs,] # overlap overlap( finds all the seqs in temp1 that also appear in temp 2)
intersectC2 <- temp4.mod[temp4.mod$OTUs %in% temp3.mod$OTUs,] # overlap overlap( finds all the seqs in temp1 that also appear in temp 2)
intersectD1 <- temp4.mod[temp4.mod$OTUs %in% temp5.mod$OTUs,] # overlap overlap( finds all the seqs in temp1 that also appear in temp 2)
intersectD2 <- temp5.mod[temp5.mod$OTUs %in% temp4.mod$OTUs,] # overlap overlap( finds all the seqs in temp1 that also appear in temp 2)


# merge the two datasets for plotting
intersectA.mod <- rbind(intersectA1, intersectA2)
intersectB.mod <- rbind(intersectB1, intersectB2)
intersectC.mod <- rbind(intersectC1, intersectC2)
intersectD.mod <- rbind(intersectD1, intersectD2)


yadjst <- 7 # so we can move around the not shared label
sampleName1 <- paste(unique(temp1.mod$Days_post)) #makes the sample name
sampleName2 <- paste(unique(temp2.mod$Days_post))
sampleName3 <- paste(unique(temp3.mod$Days_post)) #makes the sample name
sampleName4 <- paste(unique(temp4.mod$Days_post))
sampleName5 <- paste(unique(temp5.mod$Days_post))

 #legend adventures
stackedShared <- rbind(temp1.mod, temp2.mod, temp3.mod, temp4.mod, temp5.mod) #add together shared sequences
stackedShared <- stackedShared[,c(5,6,7)] #make a dataframe with just cdr3seq and fillseq number
stackedShared <- subset(stackedShared, shared!=FALSE) #remove non shared row for each sample
stackedShared = stackedShared %>% distinct #remove duplicate rows
dash <- sort(stackedShared$Genus)
dash <- as.data.frame(dash, stringsAsFactors=FALSE)
names(dash)[1] <- "Genus"
stackedShared<-merge(x=stackedShared,y=dash,by="Genus",all.x=TRUE)

ggp <- ggplot()+
geom_line(aes(x = intersectD.mod$x, y = intersectD.mod$y, group = intersectD.mod$fillseq), colour = 'grey')+
geom_line(aes(x = intersectC.mod$x + 0.9, y = intersectC.mod$y, group = intersectC.mod$fillseq), colour = 'grey')+
geom_line(aes(x = intersectB.mod$x + 1.8, y = intersectB.mod$y, group = intersectB.mod$fillseq), colour = 'grey')+
geom_line(aes(x = intersectA.mod$x + 2.7, y = intersectA.mod$y, group = intersectA.mod$fillseq), colour = 'grey')+

#geom_errorbar(aes(x = 1, ymin = round(minPcntShared-yadjst), ymax = round(minPcntShared-yadjst)))+
geom_bar(aes(x = 1, temp5.mod$Count, fill = as.factor(temp5.mod$fillseq), alpha = as.factor(temp5.mod$shared)), stat = 'identity', colour = 'black') +
  geom_bar(aes(x = 3, temp4.mod$Count, fill = as.factor(temp4.mod$fillseq), alpha = as.factor(temp4.mod$shared)), stat = 'identity', colour = 'black') + 
  geom_bar(aes(x = 5, temp3.mod$Count, fill = as.factor(temp3.mod$fillseq), alpha = as.factor(temp3.mod$shared)), stat = 'identity', colour = 'black') +
  geom_bar(aes(x = 7, temp2.mod$Count, fill = as.factor(temp2.mod$fillseq), alpha = as.factor(temp2.mod$shared)), stat = 'identity', colour = 'black') +  
  geom_bar(aes(x = 9, temp1.mod$Count, fill = as.factor(temp1.mod$fillseq), alpha = as.factor(temp1.mod$shared)), stat = 'identity', colour = 'black') + 
          scale_alpha_manual(values=c(0, 1)) +
          #geom_errorbar(aes(x = 3, ymin = round(minPcntShared-yadjst), ymax = round(minPcntShared-yadjst)))+
          scale_y_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100))+
          scale_x_continuous(breaks = c(1, 3, 5, 7, 9), labels = c(sampleName5, sampleName4, sampleName3, sampleName2, sampleName1))+
          geom_text(aes(x = 1, y = round(minPcntShared-yadjst)), label = '')+
          geom_text(aes(x = 3, y = round(minPcntShared-yadjst)), label = '')+
          geom_text(aes(x = 5, y = round(minPcntShared-yadjst)), label = '')+
          geom_text(aes(x = 7, y = round(minPcntShared-yadjst)), label = '')+
          geom_text(aes(x = 9, y = round(minPcntShared-yadjst)), label = '')+

          annotate(geom = "segment", x = -Inf, xend = -Inf, y = -Inf, yend = Inf) +
          #annotate(geom = "segment", x = -Inf, xend = -Inf, y =  round(minPcntShared-yadjst), yend = round(minPcntShared, -1),
                   #linetype = "dashed", color = "white")+ 
         # coord_cartesian(ylim=c(round(minPcntShared-yadjst), 100), expand = T) +
          labs(title = 'Shared Genera CVG017', x = ' ', y = 'Cumulative % of Sample')+
          theme(legend.position="right", legend.title=element_blank(), panel.background = element_blank()) + coord_flip() + guides(alpha = FALSE) + scale_fill_discrete(breaks=c(stackedShared$fillseq), labels = c(stackedShared$Genus))   + theme(
    plot.title = element_text(color="black", size=28, face="bold"),
    axis.title.x = element_text(color="black", size=28, face="bold"),
    axis.title.y = element_text(color="black", size=28, face="bold"),
    axis.text=element_text(size=26),
    legend.title = element_text(color = "black", size = 26),
    legend.text = element_text(color = "black", size = 20)) + guides(fill=guide_legend(ncol=1))+ labs(fill = "Genus")
        
ggsave(paste(wd, '/', 'CVG017_', paste(sampleName1, sampleName2, sampleName3, sampleName4, sampleName5, sep = '_'), '_taxa_Genus_large',  '.pdf', sep = ''), 
ggp, width = 31, height = 24, units = c('cm'))

ggp

